A Data-Driven Framework for Neural Field Modeling

### 2010-11-21 ltx/first_order_framework/LaTeX_Manuscript.tex
further general edits fixing typos

\begin{abstract}
This paper presents a framework for creating neural field models from electrophysiological data. The Wilson and Cowan or Amari style neural field equations are used to form a parametric model, where the parameters are estimated from data. To illustrate the estimation framework, data is generated using the neural field equations incorporating modelled sensors enabling a comparison between the estimated and true parameters. To facilitate state and parameter estimation, we introduce a method to reduce the continuum neural field model using a basis function decomposition to form a finite-dimensional state-space model. Spatial frequency analysis methods are introduced that systematically specify the basis function configuration required to capture the dominant characteristics of the neural field. The estimation procedure consists of a two-stage iterative algorithm incorporating the unscented Rauch-Tung-Striebel smoother for state estimation and a least squares algorithm for parameter estimation. The results show that it is theoretically possible to reconstruct the neural field and estimate intracortical connectivity structure and synaptic dynamics with the proposed framework. %The results also illustrate a systematic loss of high spatial frequency information as the cost incurred by the model reduction procedure. 
\end{abstract}

@@

We present an extension to the work of \cite{schiff2008kalman} by establishing a framework for estimating the state of the WC equations for larger scale (more space) systems via a systematic model reduction procedure. In addition, a new method is presented for estimating the connectivity structure and the synaptic dynamics. Until now, model-based estimation of local intracortical connectivity has not been reported in the literature (to the best of the authors' knowledge). Our study extends recent work which shows that it is possible to estimate local coupling of spatiotemporal systems using techniques from control systems theory and machine learning (\cite{Dewar2009}). The key development of this previous work was to represent a spatiotemporal system as a standard state-space model, with the number of states independent of the number of observations (recording electrodes in this case). In addition, the appropriate model selection tools have been developed (\cite{Scerri2009}) allowing for the application of the technique to neural fields. This paper extends the linear framework of \cite{Dewar2009} and \cite{Scerri2009} to the nonlinear case required for the neural field equations.

@@

\subsection{Derivation of Finite Dimensional State-Space Model}\label{Sect:ReducedModelDerivation}
In order to implement standard estimation techniques, we use a decomposition of the field using a set of Gaussian basis functions. Decomposition allows a continuous field to be represented by a finite-dimensional state vector. This facilitates application of standard nonlinear state estimation methods such as the unscented Kalman filter. The field decomposition is described by
\begin{equation}
	\label{DefFieldDecomp} v_t\left(\mathbf{r}\right) \approx \boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \mathbf{x}_t, 
\end{equation}
where $\mathbf{x}_t$ is the state vector that scales the field basis functions $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$. The superscript $\top$ denotes the transpose operator. The field basis functions are described by
\begin{equation}\label{eq:FieldBasisFunction}
	\phi\left(\mathbf{r}-\mathbf{r}'\right) =
\exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_{\phi}^2}\right)}, 
\end{equation}
where $\sigma_{\phi}$ is the basis function width parameter. An example of a one-dimensional field decomposition is given in Fig.~\ref{fig:Figure2}.

The width and positioning of the basis functions can be determined by spectral analysis (explained in detail in the following section). The connectivity kernel can also be decomposed as 

@@

This substitution allows us to swap the convolution in equation~\ref{reduced continuous model}, which is computationally demanding, with the convolution in equation~\ref{eq:DefPsi}, which is pre-defined analytically. This is of great importance as it lowers the computational demands of the estimation algorithm and provides a dramatic increase in estimation speed.

@@

The field basis function widths can also be inferred using spectral considerations (\cite{Sanner1992,Scerri2009}). To demonstrate this, we begin with a two-dimensional Gaussian basis function located at the origin, defined as
\begin{equation}\label{eq:BasisFunctionAtOrigin}
 \phi(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_{\phi}^2} \mathbf r^\top\mathbf r}\right).
\end{equation}
The corresponding Fourier transform is
\begin{equation}\label{eq:GaussianFT}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right),
\end{equation}
where 
\begin{equation}\label{eq:GaussianFTWidth}
	\sigma^2_{\nu} = \frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}
To obtain a 3~dB attenuation at $\boldsymbol\nu_{cy}$, the basis function width, $\sigma^2_{\phi}$, should be chosen to be
\begin{equation}\label{eq:WidthCutOffRelationship}
 \sigma^2_{\phi}= \frac{1}{\pi^2\sigma_{\nu_{cy}}^2},
\end{equation}
where
\begin{equation}\label{eq:WidthFrequencyRelationship}
 \sigma^2_{\nu_{cy}}= \frac{2\boldsymbol\nu_{cy}^\top \boldsymbol\nu_{cy}}{\ln2}.
\end{equation}
This ensures that the basis functions can represent the neural field with frequency content up to $\boldsymbol\nu_{cy}$. Derivations for equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} are given in \ref{App:FrequencyAnalysis}.

@@

The sensors were spaced on a $14 \times 14$ square grid and the sensor width was set to $\sigma^2_m = 0.81$. This provides a width of 1.5~mm at half the maximum amplitude of the observation kernel. The distance between sensors was set to $\Delta_y = 1.5$~mm. This arrangement modelled the recording as having some cross talk (from volume conduction), which is typical for neurophysiological recordings.

@@

\subsubsection{Experiment I: Monte Carlo Simulation with Fixed Parameters}
A Monte Carlo approach was employed where 150 realisations of the data were generated, such that distributions of the parameters estimates could be formed. Each realisation consisted of 500~ms of data (sampled at $1$~kHz) and the estimation was applied to the final 400~ms, allowing the model's initial transients to die out. The initial state and parameters were unknown to the estimator. Ten iterations of the estimation algorithm were used to ensure that the parameter estimates converged. Fig.~\ref{fig:Figure6} shows the rate of convergence, confirming that the algorithm converged to steady parameter values, with the changes in the parameter errors (between consecutive iterations) falling to less than $10^{-4}$ after 6 iterations.

@@

To study the robustness of the estimation algorithm, the performance was evaluated for various connectivity kernel parameters. The reconstructed connectivity kernels from the estimated parameters are shown in Fig.~\ref{fig:Figure10}. For Fig.~\ref{fig:Figure10}(A-C) the Mexican hat connectivity kernels were used. For each subset of parameters, the mean of the reconstructed kernel over 50 realisations is shown as a dotted line and the 95~\% confidence interval is indicated by the grey shaded region. In each case, the actual kernel was within the 95~\% confidence interval of the kernel estimate. The statistics for the kernel parameter estimates for each of the parameter sets are shown in Table~\ref{tab:MexicanHatKernelResults}.

For each of the Mexican hat connectivities, the support of the kernel was assumed to be known. However, in Fig.~\ref{fig:Figure10}(D-F) the kernel support in the estimator differed from the actual data. In these cases the data sets were generated using a single positive (excitatory) Gaussian kernel, as seen by the solid line in each of the subplots of the figure. The estimation algorithm was constructed allowing for the Mexican hat connectivity of the previous examples. Each plot shows that the actual kernel lies within the confidence intervals of the reconstructed estimates for the majority of the kernel space, until the amplitude is relatively low. Table~\ref{tab:GaussianKernelResults} shows the statistics of the parameter estimates. Although the net reconstructed kernel estimates are in good accordance with the actual kernels, the individual kernel basis function weights are not in agreement with the actual parameters. This result demonstrates that unless the widths of the Gaussians forming the kernel are precisely known, the interpretation of the results should be based on the net shape of the estimated connectivity kernel and not on the estimated parameter values individually.

@@

% The size of the squares in the figure show the magnitude of the log of the error. The squares are black if the log of the error is negative (the RMSE is $<1$) and the squares are white if the log of the error is positive (the RMSE is $>1$). The largest black squares indicate the best performance and the largest white squares indicate the worst performance for the parameter combinations (\dean{need some thing more quantitative}). 
From Fig.~\ref{fig:Figure11} it can be seen that the estimation accuracy depends on the parameters of the sigmoid function. This is expected since by altering the parameters we are changing the nonlinearity and the saturation region of the sigmoid. For example, when the sigmoid slope is very gentle, most of the neural activity falls in the central, nearly linear region of the sigmoid, thus having minor effects on the estimation procedure. On the other hand, sharp slopes result in most of the activity falling in the saturation region of the sigmoid with adverse effects on the estimation. Similar arguments can be made for variations to the sigmoid threshold.

@@

When the observation kernel is too narrow, high frequency components in the neural field are present but under sampled. These high frequency components result in aliasing errors in the estimated field, thus affecting the kernel estimation accuracy. On the other hand, for the very wide observation kernels considered, important neural dynamics are not pick up by the estimated field due to the filtering out of the important frequency components by the sensors, resulting in poor accuracy in the kernel estimate. As indicated in Fig.~\ref{fig:Figure12}, the best results are obtained when the sensor width is chosen to minimize both sources of errors.

@@

% {\bf Figure 12. Sigmoid functions}. From bottom left corner to the right of the second second row of the previous figure.

{\bf Figure 12. RMSE of the kernel estimates}. RMSE of the kernel for different values of sensor width, $\sigma_m$, distance between adjacent sensors is $\Delta_y=1.5$. Mean kernel estimates are obtained over 50 realisations in each case.
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig1.pdf} 
\end{center}
\caption{}
\label{fig:Figure1}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig2.pdf} 
\end{center}
\caption{}
\label{fig:Figure2}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig3.pdf}
\end{center}
\caption{} 
\label{fig:Figure3}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig4.pdf}
\end{center}
\caption{}
\label{fig:Figure4}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht] 
\begin{center}
\includegraphics{./Graph/pdf/fig5.pdf}	
\end{center}
\caption{}
\label{fig:Figure5}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig6.pdf}
\end{center}
\caption{}
\label{fig:Figure6}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig7.pdf}
\end{center}
\caption{} 
\label{fig:Figure7}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig8.pdf} 
\end{center}
\caption{} 
\label{fig:Figure8}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig9.pdf}
\end{center}
\caption{}
\label{fig:Figure9}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig10.pdf}
\end{center}
\caption{}
\label{fig:Figure10}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig11.pdf}
\end{center}
\caption{}
\label{fig:Figure11}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig12.pdf}
\end{center}
\caption{}
\label{fig:Figure12}
\end{figure*}
\clearpage
\newpage
\appendix
\section{Time Discretisation of Neural Field Model}
\label{App:Time Discretization} To form the integro-difference equation neural field model, a time discretisation must be performed. To do this, we used a one-step Euler method, where equation~\ref{FinalFormContinuous} can be approximated by 
\begin{align}
	\label{Euler Approximation}\frac{v\left( \mathbf{r},t+T_s \right) - v\left( \mathbf{r},t\right)}{T_s} &=-\zeta v\left( \mathbf{r},t \right) \nonumber \\ 
&+ \int_\Omega {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)d\mathbf{r}'}. 
\end{align}
For clarity, we shall index time points in the discrete time form of the model using the subscript $t$ and the next time point as $t+1$. Rearranging equation~\ref{Euler Approximation}, we get 
\begin{equation}
	\label{Euler Approximation2} v_{t+1}\left( \mathbf{r}\right) = v_t\left( \mathbf{r}\right) -T_s \zeta v_t\left( \mathbf{r}\right) + T_s \int_\Omega {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v_t\left( \mathbf{r}'\right)} \right)d\mathbf{r}'}. 
\end{equation}
Now making simplifications, the discrete time form of the model is 
\begin{equation}
	\label{Discrete Time Model1}v_{t+1}\left(\mathbf{r}\right) = \xi v_t\left(\mathbf{r}\right) + T_s \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'}, 
\end{equation}
where $\xi = 1 - T_s \zeta$.

\section{Observation Kernel}\label{App:ObsKernelExperiment}
The Gaussian shape of the observation kernel was verified experimentally. This was achieved by moving a dipole across a recording electrode submerged in saline. A photograph through a microscope of the submerged electrodes can be seen in \parham{Fig.~\ref{fig:SensorAndSource}}. The measurement sensor was the most central electrode of a $5\times5$ Utah array (Blackrock Microsystems) which had electrode spacing of 400~$\mu$m. The current source was a twisted wire electrode (Plastics One Inc.) with tip-tip spacing of 60~$\mu$m and tip diameter of 200~$\mu$m. The twisted wire electrode was moved in the saline solution using a 3 dimensional manipulator (Sutter Instrument Company, MPC-100) in increments of 200~$\mu$m. The twisted wire electrode and the Utah array were co-registered visually using the microscope.
The stimulation waveform for the current source consisted of a 40~Hz sine wave with a duration of 1~s. The recording from the central electrode in the array was used to estimate the shape of the sensor kernel. To extract the amplitude of the signal, the data was first band-pass filtered with a center frequency of 40~Hz. The envelope of the signal was found by computing the mean instantaneous amplitude over the 1~s stimulation duration by

@@

	a = \frac{1}{T}\sum_n{\sqrt{(y_t-\mathcal{H}y_t^2}}

@@

where $\mathcal{H}$ denotes the Hilbert transform operator, $t$ indexes samples and $T$ is the total number of samples in the time series. \parham{Fig.~\ref{fig:Ch3}} shows the mean amplitude of the recorded envelope versus the distance between the twisted wire electrode and the recording electrode. A Gaussian fitted to the amplitudes is also shown. The figure shows that a Gaussian is indeed a good model for the sensor kernel in the homogeneous solution.
\section{Spatial Cut-off Frequency of Field Basis Functions}\label{App:FrequencyAnalysis}
This section gives the derivation for the width of the Fourier transform of the field basis function in terms of its spatial cut-off frequency (equations~\ref{eq:GaussianFT} and~\ref{eq:WidthFrequencyRelationship}). We define an \textit{n}-dimensional Gaussian  centerd at the origin as 

>>>>

\begin{abstract}
This paper presents a framework for creating neural field models from electrophysiological data. The Wilson and Cowan or Amari style neural field equations are used to form a parametric model, where the parameters are estimated from data. To illustrate the estimation framework, data is generated using the neural field equations incorporating modeled sensors enabling a comparison between the estimated and true parameters. To facilitate state and parameter estimation, we introduce a method to reduce the continuum neural field model using a basis function decomposition to form a finite-dimensional state-space model. Spatial frequency analysis methods are introduced that systematically specify the basis function configuration required to capture the dominant characteristics of the neural field. The estimation procedure consists of a two-stage iterative algorithm incorporating the unscented Rauch-Tung-Striebel smoother for state estimation and a least squares algorithm for parameter estimation. The results show that it is theoretically possible to reconstruct the neural field and estimate intracortical connectivity structure and synaptic dynamics with the proposed framework. %The results also illustrate a systematic loss of high spatial frequency information as the cost incurred by the model reduction procedure. 
\end{abstract}

@@

We present an extension to the work of \cite{schiff2008kalman} by establishing a framework for estimating the state of the WC equations for larger scale (more space) systems via a systematic model reduction procedure. In addition, a new method is presented for estimating the connectivity structure and the synaptic dynamics. Until now, model-based estimation of local intracortical connectivity has not been reported in the literature (to the best of the authors' knowledge). Our study also extends recent work which shows that it is possible to estimate local coupling of spatiotemporal systems using techniques from control systems theory and machine learning (\cite{Dewar2009}). The key development of this previous work was to represent a spatiotemporal system as a standard state-space model, with the number of states independent of the number of observations (recording electrodes in this case). In addition, the appropriate model selection tools have been developed (\cite{Scerri2009}) allowing for the application of the technique to neural fields. This paper extends the linear framework of \cite{Dewar2009} and \cite{Scerri2009} to the nonlinear case required for the neural field equations.

@@

\subsection{Derivation of Finite Dimensional State-Space Model}\label{Sect:ReducedModelDerivation}
In order to implement standard estimation techniques, we use a decomposition of the field using a set of Gaussian basis functions. Decomposition allows a continuous field to be represented by a finite-dimensional state vector. This facilitates application of standard nonlinear state estimation methods such as the unscented Kalman filter. The field decomposition is described by
\begin{equation}
	\label{DefFieldDecomp} v_t\left(\mathbf{r}\right) \approx \boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \mathbf{x}_t, 
\end{equation}
where $\mathbf{x}_t$ is the state vector that scales the field basis functions $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$. The superscript $\top$ denotes the transpose operator. The field basis functions are described by
\begin{equation}\label{eq:FieldBasisFunction}
	\phi\left(\mathbf{r}-\mathbf{r}'\right) =
\exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_{\phi}^2}\right)}, 
\end{equation}
where $\sigma_{\phi}$ is the basis function width parameter. An example of a one-dimensional field decomposition is given in Fig.~\ref{fig:Figure2}. The width and positioning of the basis functions can be determined by spectral analysis (explained in detail in the following section).

The connectivity kernel can also be decomposed as 

@@

This substitution swaps the convolution in equation~\ref{reduced continuous model}, which is computationally demanding, with the convolution in equation~\ref{eq:DefPsi}, which is pre-defined analytically. This is of great importance as it lowers the computational demands of the estimation algorithm and provides a dramatic increase in estimation speed.

@@

The field basis function widths can also be inferred using spectral considerations (\cite{Sanner1992,Scerri2009}). To demonstrate this, we begin with a two-dimensional Gaussian basis function located at the origin, defined as
\begin{equation}\label{eq:BasisFunctionAtOrigin}
 \phi(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_{\phi}^2} \mathbf r^\top\mathbf r}\right).
\end{equation}
The corresponding Fourier transform is
\begin{equation}\label{eq:GaussianFT}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right),
\end{equation}
where 
\begin{equation}\label{eq:GaussianFTWidth}
	\sigma^2_{\nu} = \frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}
To obtain a 3~dB attenuation at $\boldsymbol\nu_{cy}$, the basis function width, $\sigma^2_{\phi}$, should be chosen such that
\begin{equation}\label{eq:WidthCutOffRelationship}
 \sigma^2_{\phi}= \frac{1}{\pi^2\sigma_{\nu_{cy}}^2},
\end{equation}
where
\begin{equation}\label{eq:WidthFrequencyRelationship}
 \sigma^2_{\nu_{cy}}= \frac{2\boldsymbol\nu_{cy}^\top \boldsymbol\nu_{cy}}{\ln2}.
\end{equation}
This ensures that the basis functions can represent the neural field with frequency content up to $\boldsymbol\nu_{cy}$. Derivations for equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} are given in \ref{App:FrequencyAnalysis}.

@@

The sensors were spaced on a $14 \times 14$ square grid and the sensor width was set to $\sigma^2_m = 0.81$. This provides a width of 1.5~mm at half the maximum amplitude of the observation kernel. The distance between sensors was set to $\Delta_y = 1.5$~mm. This arrangement modeled the recording as having some cross talk (from volume conduction), which is typical for neurophysiological recordings.

@@

\subsubsection{Experiment I: Monte Carlo Simulation with Fixed Parameters}
A Monte Carlo approach was employed where 150 realizations of the data were generated, such that distributions of the parameters estimates could be formed. Each realization consisted of 500~ms of data (sampled at $1$~kHz) and the estimation was applied to the final 400~ms, allowing the model's initial transients to die out. The initial state and parameters were unknown to the estimator. Ten iterations of the estimation algorithm were used to ensure that the parameter estimates converged. Fig.~\ref{fig:Figure6} shows the rate of convergence, confirming that the algorithm converged to steady parameter values, with the changes in the parameter errors (between consecutive iterations) falling to less than $10^{-4}$ after 6 iterations.

@@

To study the robustness of the estimation algorithm, the performance was evaluated for various connectivity kernel parameters. The reconstructed connectivity kernels from the estimated parameters are shown in Fig.~\ref{fig:Figure10}. For Fig.~\ref{fig:Figure10}(A-C) Mexican hat connectivity kernels were used. For each subset of parameters, the mean of the reconstructed kernel over 50 realisations is shown as a dotted line and the 95~\% confidence interval is indicated by the grey shaded region. In each case, the actual kernel was within the 95~\% confidence interval of the kernel estimate. The statistics for the kernel parameter estimates for each of the parameter sets are shown in Table~\ref{tab:MexicanHatKernelResults}.

For each of the Mexican hat connectivities, the support of the kernel was assumed to be known. However, in Fig.~\ref{fig:Figure10}(D-F) the kernel support in the estimator differed from the actual data. In these cases the data sets were generated using a single positive (excitatory) Gaussian kernel, as seen by the solid line in each of the subplots of the figure. The estimation algorithm was constructed allowing for the Mexican hat connectivity of the previous examples. Each plot shows that the actual kernel lies within the confidence intervals of the reconstructed estimates for the majority of the kernel space, until the amplitude is relatively low. Table~\ref{tab:GaussianKernelResults} shows the statistics of the parameter estimates. Although the net reconstructed kernel estimates are in good accordance with the actual kernels, the individual kernel basis function weights are not in agreement with the actual parameters. This result demonstrates that the interpretation of the results should be based on the net shape of the estimated connectivity kernel and not on the estimated parameter values individually.

@@

% The size of the squares in the figure show the magnitude of the log of the error. The squares are black if the log of the error is negative (the RMSE is $<1$) and the squares are white if the log of the error is positive (the RMSE is $>1$). The largest black squares indicate the best performance and the largest white squares indicate the worst performance for the parameter combinations (\dean{need some thing more quantitative}). 
From Fig.~\ref{fig:Figure11} it can be seen that the estimation accuracy depends on the parameters of the sigmoid function. This is expected since that by altering the parameters we are changing the nonlinearity and the saturation region of the sigmoid. For example, when the sigmoid slope is very gentle, most of the neural activity falls in the central, nearly linear region of the sigmoid, thus having minor effects on the estimation procedure. On the other hand, sharp slopes result in more of the activity falling in the saturation region of the sigmoid with adverse effects on the estimation. Similar arguments can be made for variations to the sigmoid threshold.

@@

When the observation kernel is too narrow, high frequency components in the neural field are present but under sampled. These high frequency components result in aliasing errors in the estimated field, thus affecting the kernel estimation accuracy. On the other hand, for the very wide observation kernels considered, important neural dynamics are not represented in the estimated field due to the filtering out of the important frequency components by the sensors, resulting in poor accuracy in the kernel estimate. As indicated in Fig.~\ref{fig:Figure12}, the best results are obtained when the sensor width is chosen to minimize both sources of errors.

@@

{\bf Figure 12. RMSE of the kernel estimates}. RMSE of the kernel for different values of sensor width, $\sigma_m$, distance between adjacent sensors is $\Delta_y=1.5$. Mean kernel estimates are obtained over 50 realisations in each case.
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig1.pdf} 
\end{center}
\caption{}
\label{fig:Figure1}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig2.pdf} 
\end{center}
\caption{}
\label{fig:Figure2}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig3.pdf}
\end{center}
\caption{} 
\label{fig:Figure3}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig4.pdf}
\end{center}
\caption{}
\label{fig:Figure4}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht] 
\begin{center}
\includegraphics{./Graph/pdf/fig5.pdf}	
\end{center}
\caption{}
\label{fig:Figure5}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig6.pdf}
\end{center}
\caption{}
\label{fig:Figure6}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig7.pdf}
\end{center}
\caption{} 
\label{fig:Figure7}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig8.pdf} 
\end{center}
\caption{} 
\label{fig:Figure8}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig9.pdf}
\end{center}
\caption{}
\label{fig:Figure9}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig10.pdf}
\end{center}
\caption{}
\label{fig:Figure10}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig11.pdf}
\end{center}
\caption{}
\label{fig:Figure11}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig12.pdf}
\end{center}
\caption{}
\label{fig:Figure12}
\end{figure*}

\clearpage
\newpage

\appendix
\section{Time Discretization of the Neural Field Model}
\label{App:Time Discretization} To form the integro-difference equation neural field model, a time discretization must be performed. To do this, we used a one-step Euler method, where equation~\ref{FinalFormContinuous} can be approximated by 
\begin{align}
	\label{Euler Approximation}\frac{v\left( \mathbf{r},t+T_s \right) - v\left( \mathbf{r},t\right)}{T_s} &=-\zeta v\left( \mathbf{r},t \right) \nonumber \\ 
&+ \int_\Omega {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)d\mathbf{r}'}. 
\end{align}
For clarity, we shall index time points in the discrete time form of the model using the subscript $t$ and the next time point as $t+1$. Rearranging equation~\ref{Euler Approximation}, we get 
\begin{equation}
	\label{Euler Approximation2} v_{t+1}\left( \mathbf{r}\right) = v_t\left( \mathbf{r}\right) -T_s \zeta v_t\left( \mathbf{r}\right) + T_s \int_\Omega {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v_t\left( \mathbf{r}'\right)} \right)d\mathbf{r}'}. 
\end{equation}
Now making simplifications, the discrete time form of the model is 
\begin{equation}
	\label{Discrete Time Model1}v_{t+1}\left(\mathbf{r}\right) = \xi v_t\left(\mathbf{r}\right) + T_s \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'}, 
\end{equation}
where $\xi = 1 - T_s \zeta$.

\section{Observation Kernel}\label{App:ObsKernelExperiment}
The Gaussian shape of the observation kernel was verified experimentally. This was achieved by moving a dipole across a recording electrode submerged in saline. A photograph through a microscope of the submerged electrodes can be seen in \parham{Fig.~\ref{fig:SensorAndSource}}. The measurement sensor was the most central electrode of a $5\times5$ Utah array (Blackrock Microsystems) which had electrode spacing of 400~$\mu$m. The current source was a twisted wire electrode (Plastics One Inc.) with tip-tip spacing of 60~$\mu$m and tip diameter of 200~$\mu$m. The twisted wire electrode was moved in the saline solution using a 3 dimensional manipulator (Sutter Instrument Company, MPC-100) in increments of 200~$\mu$m. The twisted wire electrode and the Utah array were co-registered visually using the microscope.
The stimulation waveform for the current source consisted of a 40~Hz sine wave with a duration of 1~s. The recording from the central electrode in the array was used to estimate the shape of the sensor kernel. To extract the amplitude of the signal, the data was first band-pass filtered with a center frequency of 40~Hz. The envelope of the signal was found by computing the mean instantaneous amplitude over the 1~s stimulation duration by
\begin{equation}
	a = \frac{1}{T}\sum_n{\sqrt{(y_t-\mathcal{H}y_t^2}}
\end{equation}
where $\mathcal{H}$ denotes the Hilbert transform operator, $t$ indexes samples and $T$ is the total number of samples in the time series. \parham{Fig.~\ref{fig:Ch3}} shows the mean amplitude of the recorded envelope versus the distance between the twisted wire electrode and the recording electrode. A Gaussian fitted to the amplitudes is also shown. The figure shows that a Gaussian is indeed a good model for the sensor kernel in the homogeneous solution.
\section{Inner Product of Two \emph{n}-Dimensional Gaussians}\label{App:InnerProdOfGaussians}
The derivation in this section was used for calculating the matrix $\mathbf \Gamma$ from equation~\ref{eq:DefGamma}. Consider two Gaussian basis functions
\begin{equation}\label{eq:n_dimensional_Gaussian1}
 \varphi_i(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_i^2} (\mathbf r-\boldsymbol \mu_i)^\top(\mathbf r-\boldsymbol \mu_i})\right)
\end{equation}
and 
\begin{equation}\label{eq:n_dimensional_Gaussian2}
\varphi_j(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_j^2} (\mathbf r-\boldsymbol \mu_j)^\top(\mathbf r-\boldsymbol \mu_j})\right).
\end{equation}

The inner product of the two \emph{n}-dimensional Gaussians is defined as
\varphi_i\otimes\varphi_j \triangleq \int_{\mathbb{R}^n}\varphi_i(\mathbf r)\varphi_j(\mathbf r)d\mathbf{r}.
A Gaussian basis function is symmetric with respect to its center, therefore
Now we solve the convolution using the formula derived in \ref{ap:ConvOfGaussians}, and evaluating it at $\mathbf r=2\boldsymbol\mu_i $ gives
In this section, we provide a derivation for the convolution of two $n$-dimensional Gaussian basis functions. This derivation is used in the analytic calculation of $\boldsymbol\Psi(\mathbf{r})$ and the elements of the observation matrix, $\mathbf C$. Consider the two Gaussian basis functions from equations~\ref{eq:n_dimensional_Gaussian1} and~\ref{eq:n_dimensional_Gaussian2}. The derivation proceeds by assuming the Gaussian basis functions are positioned at the origin ($\boldsymbol\mu_i=\boldsymbol\mu_j=\mathbf 0$). The convolution theorem states that the Fourier transform of the convolution of two functions is equal to the product of their Fourier transforms. Using the formula from the derivation of the Fourier transform of a \emph{n}-dimensional Gaussian from equation~\ref{eq:FTofGaussian} in \ref{App:FrequencyAnalysis}, we have
\section{Mean and Covariance of the State Disturbance }\label{App:ColoredNoise} 
In this section, we show that the state disturbance vector is $\mathbf{e}_t\sim\mathcal{N}(\mathbf 0,\boldsymbol\Sigma_e)$ if $e_t(\mathbf{r})\sim\mathcal{GP}\left(\mathbf 0,\gamma(\mathbf{r}-\mathbf{r}')\right)$. We begin the derivation by acknowledging the state disturbance vector is a linear function of $e_t(\mathbf r)$. Therefore, $\mathbf{e}_t$ also follows a Gaussian distribution. From equation~\ref{eq:Wt}, the state disturbance vector is defined as

@@

	\mathbf{e}_t \triangleq \boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi} ( \mathbf{r} )e_t( \mathbf{r} )d\mathbf{r}}.

@@

The expected value of $\mathbf e_t$ is given by 
	\mathbf E\left[ \mathbf e_t\right]= \mathbf{\Gamma}^{-1}\int_{\Omega}\boldsymbol\phi\left(\mathbf{r}\right)\mathbf E\left[e_t\left(\mathbf{r}\right)\right] d\mathbf{r}=\mathbf 0.
The covariance of $\mathbf{e}_t$ is 
	\boldsymbol\Sigma_e &=\lefteqn{\mathbf E\left[ \mathbf e_t\mathbf e_t^\top\right]} \nonumber \\ 
&=\mathbf{\Gamma}^{-1}\mathbf E[\int_{\Omega}\boldsymbol{\phi}(\mathbf{r})e_t(\mathbf{r})d\mathbf{r} \int_{\Omega}\boldsymbol{\phi}^{\top}(\mathbf{r}') e_t(\mathbf{r}')d\mathbf{r}']\mathbf{\Gamma}^{- \top} \nonumber \\
	&=\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega} \boldsymbol{\phi}(\mathbf{r}) \mathbf E[e_t(\mathbf{r})e_t(\mathbf{r}')]\boldsymbol{\phi}^{\top}(\mathbf{r}')d\mathbf{r}' d\mathbf r\mathbf{\Gamma}^{- \top} \nonumber\\
	&=\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}(\mathbf r) \gamma(\mathbf r- \mathbf r' )\boldsymbol{\phi}^{\top}(\mathbf r')d\mathbf r' d\mathbf r\mathbf{\Gamma}^{-\top}.
